\chapter{Hilbert symbol}

\section{Hilbert symbol over a field}

\begin{definition}
	\label{def:Hilbert_symbol}
	\lean{hilbertSym}
	\leanok
	Let $k$ be a field and $a, b \in k$. If both $a,b$ are nonzero, the \emph{Hilbert symbol} $(a, b)_k$ is defined as
	follows: if there exist $x, y, z \in k$, not all zero, such that $z^2 = ax^2 + by^2$,
	then $(a, b)_k = 1$; otherwise, $(a, b)_k = -1$. If one of $a,b$ is zero, we set $(a, b)_k = 0$.
\end{definition}


\subsection{Basic properties of the Hilbert symbol}

The Hilbert symbol has the following properties, useful for computing it in practice.

\begin{lemma}
	\label{lem:well_defined_up_to_squares}
	\lean{hilbertSym.mul_square_eq}
	\uses{def:Hilbert_symbol}
	\leanok
Let $a, b, a', b' \in k^\times$. Then $(a a'^2, b b'^2)_k = (a,b)_k$.
\end{lemma}

\begin{proof}
	\leanok
	Let $x, y, z \in k$ be such that $z^2 = a a'^2 x^2 + b b'^2 y^2$. Then, $xa', yb', z$ are not all
	zero and satisfy $z^2 = ax^2 + by^2$. Conversely, if $x, y, z \in k$ are such that
	$z^2 = ax^2 + by^2$, then $x/a', y/b', z$ are not all zero and satisfy
	$z^2 = a a'^2 x^2 + b b'^2 y^2$.
\end{proof}

The following lemma uses the definition \code{QuadraticAlgebra} from \mathlib.

\begin{lemma}
	\label{lem:eq_one_iff}
	\lean{hilbertSym.eq_one_iff}
	\uses{def:Hilbert_symbol}
	\leanok
	Let $a, b \in k^\times$. Then $(a, b)_k = 1$ if and only if there exists $t \in k(\sqrt{b})$ such
	that $a = N_{k(\sqrt{b})/k}(t)$.
\end{lemma}

\begin{proof}
	\leanok
	\begin{itemize}
		\item Case $b$ is a square in $k$. Then $k(\sqrt{b}) = k$, and the norm map is the identity, so
		$a$ is always the norm of itself. Let $c \in k^\times$ be such that $c^2 = b$.
		Then $c^2 = a \cdot 0^2 + b \cdot 1^2$, so $(a,b)_k=1$.
		\item Case $b$ is not a square in $k$. Then $k(\sqrt{b})$ is a quadratic extension of $k$, and
		the norm map is given by $N_{k(\sqrt{b})/k}(p + q \sqrt{b}) = p^2 - b q^2$.
		Suppose $a$ is a norm, so it satisfies $a = p^2 - b q^2$ for some $p, q \in k$.
		Then $1, q, p$ are not all zero and satisfy $ p^2 = a \cdot 1^2 + b \cdot q^2$,
		so $(a, b)_k = 1$. Conversely, if $x, y, z$ are not all zero and satisfy $z^2 = ax^2 + by^2$,
		then $x=0$, since otherwise $b$ would be a square in $k$, which contradicts our assumption.
		Then $x \ne 0$, and we have $a = (z/x)^2 - b (y/x)^2$, so $a$ is a norm.
	\end{itemize}
\end{proof}

\begin{proposition}
	\label{prop:hilbert_symbol_properties}
	\lean{hilbertSym.comm, hilbertSym.right_square_eq_one, hilbertSym.right_neg_self_eq_one,
	hilbertSym.right_one_minus_self_eq_one, hilbertSym.right_mul_eq_of_eq_one , hilbertSym.right_neg_mul,
	hilbertSym.right_minus_self_mul}
	\uses{def:Hilbert_symbol}
	\leanok
	The Hilbert symbol statisfies the following formulas:
	\begin{enumerate}[(i)]
		\item For $a, b \in k$, we have $(a, b)_k = (b, a)_k$.
		\item Let $a, c \in k^\times$. Then $(a, c^2)_k = 1$.
		\item For $a \in k^\times$, we have $(a, -a)_k = 1$.
		\item Let $a \in k$ with $a \ne 0, 1$. Then $(a, 1 - a)_k = 1$.
		\item Let $a, a', b \in k^\times$ and assume that $(a, b)_k = 1$. Then $(a a', b)_k = (a', b)_k$.
		\item Let $a, b \in k$. Then $(a, -ab)_k = (a, b)_k$.
		\item Let $a, b \in k^\times$ with $a \ne 1$. Then $(a, (1-a)b)_k = (a, b)_k$.
	\end{enumerate}
\end{proposition}

\begin{proof}
	\uses{def:Hilbert_symbol, lem:eq_one_iff, lem:well_defined_up_to_squares}

	\begin{itemize}
		\item Trivial if $a$ or $b$ are zero. Otherwise, if $x,y,z$ is a solution to $z^2 = ax^2 + by^2$, then $y,x,z$ is a solution to
	$z^2 = bx^2 + ay^2$.
		\item In this case, $c^2 = a \cdot 0^2 + b \cdot 1^2$, so $(a,b)_k=1$.
		N.B. Alternatively: the same argument as in the first case of the proof of Lemma~\ref{lem:eq_one_iff} applies.
		\item In this case, $0 = a \cdot 1^2 + (-a) \cdot 1^2$.
		\item In this case, $1 = a \cdot 1^2 + (1 - a) \cdot 1^2$.
		\item We know, by Lemma~\ref{lem:eq_one_iff}, that there exists $t \in k(\sqrt{b})$ such that
		$a = N_{k(\sqrt{b})/k}(t)$. If $(a',b)_k=1$, then $a'$ is the norm of some $t' \in k(\sqrt{b})$,
		so $aa' = N_{k(\sqrt{b})/k}(t t')$, so $(aa', b)_k = 1$. Conversely, if $(aa', b)_k = 1$, then
		$aa' = N_{k(\sqrt{b})/k}(t'')$ for some $t'' \in k(\sqrt{b})$, so
		$a' = N_{k(\sqrt{b})/k}(t''/t)$, hence $(a', b)_k = 1$.
		\item Trivial if $a$ or $b$ are zero. Otherwise it follows from the previous steps, since $(a, -a)_k = 1$.
		\item Follows from the previous steps, since $(a, 1 - a)_k = 1$.
	\end{itemize}
\end{proof}
\subsection{Local Properties}
We now specialize to the case where $k$ is either $\R$ or $\Q_p$ for some prime $p$. In this case, the Hilbert symbol can be computed explicitly.

\begin{theorem}
	\label{thm:real}
	\lean{hilbertSym.real_eq}
	\uses{def:Hilbert_symbol}
	\leanok
	Let $a, b \in \R^\times$. Then $(a, b)_{\R} = 1$ if and only if $a > 0$ or $b > 0$; otherwise $(a, b)_{\R} = -1$.
\end{theorem}

\begin{proof}
	If $a>0$, then $(1,0,\sqrt{a})$ works. Similarly, if $b>0$, then $(0,1,\sqrt{b})$ works.
	If both $a$ and $b$ are negative, then $ax^2 + by^2 \le 0$ for all $x,y$, so there is no
	nontrivial solution to $z^2 = ax^2 + by^2$.
\end{proof}

The following uses the \mathlib API for $p$-adic numbers and integers, as well as the results from the files \code{HassePrinciple/Padics/Lemmas.lean} and \code{HassePrinciple/Padics/Legendre.lean}.


For the statement of the theorem in the $p$-adic case, we need the following definitions.


\begin{definition}
	\label{def:epsilon}
	\lean{PadicInt.epsilon}
	\leanok
	Let $u \in (\Z_2)^\times$. We define $\epsilon(u)$ to be the class modulo $2$ of $\dfrac{u-1}{2}$.
\end{definition}

\begin{definition}
	\label{def:omega}
	\lean{PadicInt.omega}
	\leanok
	Let $u \in (\Z_2)^\times$. We define $\omega(u)$ to be the class modulo $2$ of $\dfrac{u^2-1}{8}$.
\end{definition}



For the proof, we will need the following lemma.
\begin{lemma}
	\label{lem:nontrivialzero}
	\lean{Padic.exists_nontrivial_zero}
	\leanok
Let $p$ be a prime and let $v \in (\Q_p)^\times$. Let $x, y, z \in \Q_p$ be such that
$(x, y, z) \ne (0, 0, 0)$ and $z^2 - p x^2 - v y^2 = 0$. Then there exist
$z', y' \in (\Z_p)^\times$ and $x' \in \Z_p$ such that $(z', y', x')$ is a nontrivial solution to
the same equation.
\end{lemma}
\begin{proof}
	Let $(x,y,z)$ be any nontrivial solution. For any $\lambda \ne 0$, also $(\lambda x, \lambda y,
	\lambda z)$ is a solution. Taking $\lambda = p^a$ for an appropriate $a$, we can assume that
	$x,y,z$ are all in $\Z_p$ and at least one of them is a unit. Now, if $z$ is not a unit, then
	$px^2+vy^2$ is also not a unit, thus $y$ is not a unit. Conversely if $y$ is not a unit, then
	$z^2-px^2$ is also not a unit, thus $z$ is not a unit. This shows that if eithe $y$ or $z$ is not
	a unit, then they are both non units, but then $p^2$ would divide $px^2$, hence $x$ would not be a
	unit either, contradicting the fact that one of $x,y,z$ is a unit.
\end{proof}

We are now ready to compute the Hilbert symbol $(x,y)$ in the $p$-adic case.
We split the statement and the proof of the theorem for the $p$-adic into the cases where $p$ is odd
and where $p = 2$, and for each we first consider separately the cases where the valuations of $x$
and $y$ are both $0$, both $1$, or one of them is $0$ and the other is $1$.

Everywhere in the following, we assume that $x, y \in (\Q_p)^\times$, and denote by $v_x$ and $v_y$
their valuations, and by $u_x$ and $u_y$ their unit parts.

\begin{theorem}
	\label{thm:padic_odd_p}
	\lean{hilbertSym.padic_odd_eq, hilbertSym.padic_odd_case00, hilbertSym.padic_odd_case10, hilbertSym.padic_odd_case11}
	\uses{def:Hilbert_symbol, def:padicUnitPart, def:epsilon, def:legendreSym}
	\leanok
Let $p$ be an odd prime and let $x,y \in (\Q_p)^\times$. Then
\[(x,y) = (-1)^{v_x v_y\epsilon(p)}\left(\dfrac{u_x}{p}\right) \left(\dfrac{u_y}{p}\right),\]
where $\left(\dfrac{\cdot}{p}\right)$ is the Legendre symbol.
\end{theorem}

\begin{proof}
	\uses{lem:nontrivialzero, prop:hilbert_symbol_properties, lem:multivariable_hensel}
	Since the Hilbert symbol is well defined up to squares, we can assume that $v_x, v_y \in \{0, 1\}$.
	\begin{itemize}
		\item Suppose $v_x=v_y=0$. Then $x=u_x$ and $y=u_y$ are units. By the theorem of Chevalley--Warning,
		the equation $z^2 = u_x x^2 + u_y y^2$ has a nontrivial solution in $\F_p$, so it has a
		nontrivial solution in $\Z_p$ (by multivariate Hensel's lemma), so $(x,y)=1$.
		Note: Chevalley--Warning is in \mathlib as \code{char_dvd_card_solutions}, in \code{Mathlib.FieldTheory.ChevalleyWarning}.
		\item Suppose $v_x=1$ and $v_y=0$. Then $x=p u_x$ and $y=u_y$. Since by the previous case, we
		have $(u_x, u_y)=1$, we have $(x,y) = (p u_x, u_y) = (p, u_y)$. We just need to show that this is
		equal to $\left(\dfrac{u_y}{p}\right)$. Observe that a unit is a square in $\Q_p$ if and only if
		it is a square modulo $p$. Now, if $u_y$ is a square, then both terms are $1$, see
		\ref{prop:hilbert_symbol_properties}. Otherwise, both terms are $-1$. In fact,
		if $z^2-p^2x^2-u_yy^2$ has a nontrivial solution, we can assume that $z$ and $u_y$ are units, so
		we can reduce modulo $p$ to get $u_y = (zy^{-1})^2$, which is a contradiction.
		\item Suppose $v_x=v_y=1$. Then $x=p u_x$ and $y=p u_y$. This case follows from the properties of
		the Hilbert symbol and the previous case: $(pu_x,pu_y)=(pu_x, -p^2u_xu_y)= (pu_x, -u_xu_y)=
		\left(\dfrac{-u_xu_y}{p}\right)$. By the properties of the Legendre symbol, this is equal to
		$\left(\dfrac{-1}{p}\right)\left(\dfrac{u_x}{p}\right)\left(\dfrac{u_y}{p}\right)$, as claimed.
	\end{itemize}
\end{proof}

We proceed similarly for the case $p = 2$.

\begin{theorem}
	\label{thm:padic_2}
	\lean{hilbertSym.two_adic_eq, hilbertSym.two_adic_case00, hilbertSym.two_adic_case10, hilbertSym.two_adic_case11}
	\uses{def:Hilbert_symbol, def:padicUnitPart, def:epsilon, def:omega, def:legendreSym}
	\leanok
Let $x,y \in (\Q_2)^\times$. Then
\[(x,y) = (-1)^{\epsilon(u_x)*\epsilon(u_y) + v_x * \omega(u_y) + v_y * \omega(u_x)}.\]
\end{theorem}

\begin{proof}
	\uses{prop:hilbert_symbol_properties, lem:nontrivialzero, lem:multivariable_hensel}
	Similarly to the previous case, we can assume that $v_x, v_y \in \{0, 1\}$.

	\noindent\textbf{Case 1} Suppose $v_x=v_y=0$. Then $x=u_x$ and $y=u_y$ are units. The claim is
		$(u_x,u_y)=(-1)^{\epsilon(u_x)\epsilon(u_y)}$, which is $1$ exactly when at least one of $u_x$
		and $u_y$ is congruent to $1$ modulo $4$ and $-1$ otherwise.
		Suppose first, without loss of generality, that $u_x \equiv 1 \pmod 4$. Then its residue class
		modulo $8$ is either $1$ or $5$. In the first case, $u_x$ is a square, so $(u_x,u_y)=1$. In the
		second case, $u_x+4u_y \equiv 1 \pmod 8$, so it is the square of some $w in \Z_2$, so
		\[
		w^2-u_x\cdot 1^2- u_y\cdot 2^2 = 0
		\]
		and thus $(u_x,u_y)=1$.

		If $u_x \equiv u_y \equiv -1 \pmod 4$, then suppose there exist (X,Y,Z) not all zero, coprime,
		such that
		\[
		Z^2-u_xX^2-u_yY^2=0.
		\]
		Reducing modulo $4$, we obtain $Z^2+X^2+Y^2 \equiv 0 \pmod 4$, which is impossible, since the
		only squares modulo $4$ are $0$ and $1$, and the sum can be $0$ only if $X,Y,Z$ are all even,
		contradicting the fact that they are coprime. Thus, $(u_x,u_y)=-1$.
		This proof uses \code{ZMod.exists_sq_eq_two_iff} from \code{Mathlib.NumberTheory.LegendreSymbol.QuadraticReciprocity} and \code{hensels_lemma} in \code{Mathlib.NumberTheory.Padics.Hensel}

		\noindent\textbf{Case 2} Suppose $v_x=1$ and $v_y=0$. Then $x=2 u_x$ and $y=u_y$. We first show that
		\[
		(2, u_y) = (-1)^{\omega(u_y)},
		\]
		i.e. that $(2, u_y)=1$ if an only if $u_y \equiv \pm 1 \pmod 8$.
		Suppose that $(2, u_y)=1$. Equivalently, there exist $X,Y,Z$ not all zero, coprime, such that
		\[
		Z^2-2X^2-u_yY^2=0.
		\]
		By Lemma \ref{lem:nontrivialzero}, we can assume that $Z$ and $Y$ are units.
		Reducing modulo $8$, we get $1 - 2X^2 - u_y \equiv 0 \pmod 8$. Since squares modulo $8$ are
		$0, 1$ and $4$, we have $2X^2 \equiv 0$ or $2 \pmod 8$, so $u_y \equiv 1$ or $-1 \pmod 8$.

		Conversely, if $u_y \equiv 1 \pmod 8$, then it is a square, so $(2, u_y)=1$.
		If $u_y \equiv -1 \pmod 8$, then $Z^2-2X^2-u_yY^2=0$ has a solution modulo $8$, namely $(1,1,1)$,
		which can be lifted to a solution in $\Z_2$ by Hensel's lemma (\ref{lem:multivariable_hensel}),
		so $(2, u_y)=1$. Now, if we prove that $(2u_x,u_y)=(u_x,u_y)(2,u_y)$, then we are done. If either
		factor is $1$, by Proposition \ref{prop:hilbert_symbol_properties}, we are done. Otherwise, both
		factors are $-1$, so by the above we know $u_x \equiv -1 \pmod 4$ and $u_y \equiv 3 \pmod 8$. Up
		to multiplying by squares, we can consider the following two cases:
		\begin{itemize}
			\item $u_x=-1,u_y=3$. Then $(1,1,1)$ is a nontrivial solution to $Z^2+2X^2-3Y^2=0$.
			\item $u_x=3,u_y=-5$. Then $(1,1,1)$ is a nontrivial solution to $Z^2-6X^2+5Y^2=0$.
		\end{itemize}

		\noindent\textbf{Case 3} Suppose $v_x=v_y=1$. Using Proposition \ref{prop:hilbert_symbol_properties} and the proof
		of the previous case, we have
		\[
		(2u_x,2u_y)=(2u_x,-4u_xu_y)=(2u_x,-u_xu_y)=(-1)^{\epsilon(u_x)\epsilon(-u_xu_y)+\omega(-u_xu_y)},
		\]
		and a direct calculation shows that
		\[
		\epsilon(u_x)\epsilon(-u_xu_y)+\omega(-u_xu_y) = \epsilon(u_x)\epsilon(u_y) + \omega(u_y) + \omega(u_x).
		\]
\end{proof}

\subsection{Global properties of the Hilbert symbol}

This section uses the \mathlib function \code{Nat.nth Nat.Prime}. The main goal is to prove
the product formula for the Hilbert symbol, which states that for every $a, b \in \Q^\times$, we have
\[\prod_{v \in V} (a, b)_v = 1.\]
Here, $V$ is the union of the set of prime numbers and the symbol $\infty$, and for each prime $p$,
the symbol $(a, b)_p$ is the Hilbert symbol computed in $\Q_p$, while $(a, b)_\infty$ is the Hilbert
symbol computed in $\R$.


\begin{theorem}
	\label{thm:almost_all_one}
	\lean{hilbertSym.almost_all_one}
	\uses{def:Hilbert_symbol}
	\leanok
Let $a, b \in \Q^\times$. Then $(a, b)_p = 1$ for all but finitely many primes $p$.
\end{theorem}

\begin{proof}
	From the explicit formulas for the Hilbert symbol in the $p$-adic and in the real case, we see
	that it is multiplicative in both variables. 	In particular, it suffices to prove this theorem
	(and the following one) for $a$ and $b$ equal to $-1$ or
	a prime number. Using the formulas, we find:
	\begin{itemize}
	\item Case $a = -1, \, b = -1$. One has $(-1, -1)_\infty = (-1, -1)_2 = -1$ and
	$(-1, -1)_p = 1$ if $p \ne 2, \infty$;
  \item $a = -1, \, b = \ell$ with $\ell$ prime. If $\ell=2$, one has $(-1, 2)_v = 1$
	for all places $v$; if $v \ne 2$, one has $(-1, \ell)_v = 1$ if $v \ne 2, \ell$,
	and $(-1, \ell)_2 = (-1, \ell)_{\ell} = (-1)^{e(\ell)}$;
  \item $a = \ell, b = \ell'$ with $\ell, \ell'$ primes. If $\ell = \ell'$, by Proposition
	\ref{prop:hilbert_symbol_properties}, $(\ell, \ell)_v = (-1, \ell)_v$ for all $v$ and we are
	reduced to the case considered above. If $\ell \ne \ell'$ and if $\ell' = 2$, one has
	$(\ell, 2)_v = 1$ for $v \ne 2, \, \ell$ and
\[
(\ell, 2)_2 = (-1)^{\omega(\ell)}, \qquad (\ell, 2)_{\ell} = \left(\frac{2}{\ell}\right) = (-1)^{\omega(\ell)}, \quad \text{cf. chap. I, n° 3.2, th. 5.}
\]
If \( \ell \) and \( \ell' \) are distinct and different from 2, one has \( (\ell, \ell')_v = 1 \) for \( v \neq 2, \, \ell, \, \ell' \) and
\[
(\ell, \ell')_2 = (-1)^{\epsilon(\ell)e(\ell')}, \quad (\ell, \ell')_{\ell} = \left(\frac{\ell'}{\ell}\right), \quad (\ell, \ell')_{\ell'} = \left(\frac{\ell}{\ell'}\right).
\]
\end{itemize}
\end{proof}

\begin{theorem}
	\label{thm:product_formula}
	\lean{hilbertSym.prod_eq_one}
	\uses{def:Hilbert_symbol}
	\leanok
Let $a, b \in \Q^\times$. Then $\prod_{v \in V} (a, b)_v = 1$.
\end{theorem}
\begin{proof}
	\uses{thm:almost_all_one, thm:real, thm:padic_odd_p, thm:padic_2, prop:hilbert_symbol_properties}
	Follows immediately from the proof of Theorem \ref{thm:almost_all_one} and quadratic reciprocity,
	which can be imported from \code{Mathlib.NumberTheory.LegendreSymbol.QuadraticReciprocity}.
	In all cases, the product is one.
\end{proof}

\subsection{Approximation and Existence Theorem}

\begin{definition}
	\label{def:finite_embedding}
	\lean{Rat.finiteEmbedding}
	\leanok
Let $S$ be a finite set of prime numbers.
We define the \emph{finite embedding} of $\Q$ into $\R \times \prod_{p \in S} \Q_p$ as the diagonal map.

N.B. more precisely, in the Lean code, S is a subset of the natural numbers, and the product is
over the $n$-th prime for $n \in S$.
\end{definition}

Note: we will care about the product of finitely many completions of $\Q$, which do not necessarily
include $\R$, but the proof of the Approximation Theorem assumes it without loss of generality,
since we can always add $\infty$ to $S$.

\begin{theorem}\label{thm:approximation}%todo change references
	\lean{Rat.approximation}
	\uses{def:finite_embedding}
	\leanok
	Let $S$ be a finite set of prime numbers.
	The image of the finite embedding of $\Q$ into $\R \times \prod_{p \in S} \Q_p$ is dense with respect to
	the product topology. More concretely, for every $\epsilon > 0$ and every
	$y \in \R \times \prod_{p \in S} \Q_p$, there exists $x \in \Q$ such that the distance
	between $y$ and the image of $x$ is less than $\epsilon$.
\end{theorem}
\begin{proof}
	Let $y=(y_\infty, (y_p)_{p \in S})$ be an element of $\R \times \prod_{p \in S} \Q_p$.
	We can assume $y_p \in \Z_p$ for all $p \in S$, since otherwise we can multiply all $y_p$'s by an
	integer. Let $\varepsilon > 0$ and $N \ge 0$. We need to prove that there exists $x \in \Q$ such that
	$|y_\infty - x| \le \varepsilon$ and $v_p(y_p - x) \ge N$ for all $p \in S$.

	By the Chinese Remainder Theorem \code{Nat.chineseRemainderOfFinset}, there exists $x_0 \in \N$
	such that $x_0 \equiv y_p \pmod{p^N}$ for all $p \in S$.

	Fix $q$ a prime number. By density of the numbers of the form $a/q^m$ in $\R$, where $a \in \Z$
	and $m \ge 0$, we can find $a$ in $\Z$, $q$ a prime not in $S$ and $m \ge 0$, such that
	\[
	|x_0 - x_\infty - \frac{a}{q^m}(p_1\dots p_n)^N| \le \varepsilon.
	\]

	Then $x = x_0 + \frac{a}{q^m}(p_1\dots p_n)^N$ satisfies the required conditions.
\end{proof}


\begin{theorem}\label{thm:existence}
	\lean{hilbertSym.exists_rat_with_finite_prescribed_hilbertSym}
	\uses{def:Hilbert_symbol}
	\leanok
	Let $(a_i)_{i \in I}$ be a finite family of nonzero rational numbers, and let
	$e_{i,v} \in \{\pm 1\}$ for each $i \in I$ and each place $v$ of $\Q$. There exists a rational
	number $x \in \Q^\times$ such that
	\[
	\forall i \in I, \forall v \text{ place of } \Q, (x,a_i)_v=e_{i,v}
	\]
	if and only if the following conditions hold:
	\begin{enumerate}
		\item For each $i \in I$, we have $e_{i,v}=1$ for all but finitely many places $v$.
		\item For each $i \in I$, we have $\prod_v e_{i,v}=1$.
		\item For each place $v$, there exists $x_v \in \Q_v^\times$ such that
		$(x_v, a_i)_v = e_{i,v}$ for all $i \in I$.
	\end{enumerate}
\end{theorem}

The proof uses \code{Nat.infinite_setOf_prime_and_modEq}.

\begin{proof}
	\uses{def:Hilbert_symbol, thm:approximation, thm:real, thm:padic_odd_p, thm:padic_2,
	prop:hilbert_symbol_properties, thm:product_formula, thm:padic_squares_open}
To show that the conditions are necesary, simply observe that 1. is Theorem \ref{thm:almost_all_one},
2. is Theorem \ref{thm:product_formula}, and 3. is trivial since we can take $x_v = x$ for all $v$.

Conversely, suppose that 1., 2.\ and 3.\ hold. Up to multiplying by a square integer (which does
not change the Hilbert symbols), we can assume that all $a_i$'s are integers.
Let $S$ be the set of places $2$, $\infty$ and all the prime divisors of the $a_i$'s.
Let $T$ be the set of places $v$ such that there exists $i \in I$ with $e_{i,v}=-1$.
The set $S$ is finite and, by 1., $T$ is also finite.

Consider first the case where $S \cap T= \emptyset$. Then, we define
\[
a = \prod_{\ell \in T} \ell, \quad m = 8 \prod_{\ell \in S \setminus\{2,\infty\}} \ell.
\]

Since $a,m$ are coprime, by \code{Nat.infinite_setOf_prime_and_modEq}, there exists a prime number
$p$ such that $p \equiv a \pmod m$ and $p \not\in S \cup T$.

Taking $x=ap$ works.
\begin{itemize}
	\item If $v= \infty$, then $x > 0$, so $(x, a_i)_\infty = 1$ for all $i$, and $\infty \notin T$,
	so $e_{i,\infty}=1$.
	\item If $v = \ell \in S$, we have $x \equiv a^2 \pmod m$, so $x$ is a square modulo any $\ell \in S$,
	hence by Hensel's lemma it is a square in $\Q_\ell$, and $x$ is a square modulo $8$, so again by
	Hensel's lemma it is a square in $\Q_2$. By Proposition \ref{prop:hilbert_symbol_properties},
	$(x,a_i)=1$.
	\item If $v = \ell \notin T$ and $\ell \ne p$, $v_\ell(x)=0$ and so
	$(x,a_i)_{\ell}=\left(\dfrac{a_i}{\ell}\right)^v_{\ell}(x)=1$. On the other hand, $e_{i,\ell=1}$,
	since $v \notin T$.
	\item If $v = \ell in T$, then $v_\ell(x)=1$, so $(x,a_i)_{\ell}=\left(\dfrac{a_i}{\ell}\right)$.
  By 3., there exists $x_\ell \in \Q_\ell^\times$ such that $(x_\ell, a_i)_\ell = e_{i,\ell}$ for all $i$,
	and $v_\ell(x_\ell)$ is odd since at least one of the $e_{i,\ell}$'s is $-1$.
	Therefore, $(a_i,x)_\ell$ and $(a_i,x_\ell)_\ell$ are both equal to $\left(\dfrac{a_i}{\ell}\right)$.
	\item If $v = p$, since for all the other places the equality holds, by the product formula we
	also have $(x,a_i)_p = e_{i,p}$ for all $i$.
\end{itemize}

For the general case, since the image of $\Q^\times$ is dense in $\prod_{v \in S} \Q_v^\times$, by
Theorem \ref{thm:approximation}, and squares are open in $\Q_v^\times$, by Theorem \ref{thm:padic_squares_open},
we can find $x' \in \Q^\times$ such that $x'/x_v$ is a square in $\Q_v^\times$ for all $v \in S$.
Then, $(x', a_i)_v = (x_v, a_i)_v = e_{i,v}$ for all $i$ and all $v \in S$.
Let $\eta_{i,v}= e_{i,v} (x', a_i)_v$. Then $\eta_{i,v}=1$ for all $v \in S$ and conditions 1., 2.\
and 3. hold for the $\eta_{i,v}$'s, so we can apply the first part of the proof to find $y \in \Q^\times$
such that $(y, a_i)_v = \eta_{i,v}$ for all $i$ and all $v$.

Finally, setting $x = x' y$, we have $(x, a_i)_v = (x', a_i)_v (y, a_i)_v = e_{i,v}$ (by the explicit
formulas for the Hilbert symbol in the $p$-adic and in the real case) for all $i$ and all $v$, as desired.
\end{proof}
